CF Registry Analysis

### 2023-12-19 Christine Chan/CF Registry/cf_registry_analysis.qmd
reorganize home collection code

# Survival analysis

@@

## Participants with a single observation

```{r}
# Get total number of people with only one row with CFRD info
one_row <- annual %>%
  group_by(eDWID) %>%
  summarise(n = n()) %>%
  filter(n == 1)
one_row_cfrd <- annual %>%
  filter(!is.na(cfrd_status_annual)) %>%
  group_by(eDWID) %>%
  summarise(n = n()) %>%
  filter(n == 1)
```

There were `r nrow(one_row)` people with a single year of observation in the annual data (across all variables), and `r nrow(one_row_cfrd)` people with a single non-missing CFRD status. `r length(intersect(one_row$eDWID,one_row_cfrd$eDWID))` people fall into both of these categories:

```{r}
plot(
  euler(list(
    "One Observation (All)" = one_row$eDWID,
    "One Observation (CFRD)" = one_row_cfrd$eDWID
  )),
  fills = c("lightblue2", "lightsalmon"),
  quantities = T,
  legend = list(side = "right")
)
```

## Participants with $\geq$ 3 years between CFRD+

```{r}
# I can't think of a good tidyverse way to do this, so trying it myself
dfs <- annual %>% select(eDWID, ReviewYear, cfrd_status_annual)
dfs <- split.data.frame(dfs, dfs$eDWID)
# For each ID, test whether they follow the pattern
fit_pattern <- lapply(dfs, function(d) {
  r <- rle(as.character(d$cfrd_status_annual))
  # Check for CFRD+, CFRD-, CFRD+ pattern
  # Can have NAs on either side of the CFRD- as well, so
  # CFRD+, CFRD-, CFRD+ OR CFRD+, NA, CFRD-, CFRD+ OR CFRD+, CFRD-, NA, CFRD+
  # OR CFRD+, NA,CFRD-, NA,CFRD+
  if (
    grepl(
      paste("CFRD\\+,CFRD\\-,CFRD\\+|", "CFRD\\+,NA,CFRD\\-,CFRD\\+|",
        "CFRD\\+,CFRD\\-,NA,CFRD\\+|", "CFRD\\+,NA,CFRD\\-,NA,CFRD\\+",
        sep = "", collapse = ""
      ),
      paste(r$values, collapse = ",")
    )) {
    return(unique(d$eDWID))
  } else {
    return(NA)
  }
})
fit_pattern <- fit_pattern[!is.na(fit_pattern)]
# Filter out those that don't follow the pattern, calculate largest gaps
# between CFRD+ diagnoses
largest_gap <- annual %>%
  filter(eDWID %in% names(fit_pattern), cfrd_status_annual == "CFRD+") %>%
  select(eDWID, ReviewYear, cfrd_status_annual) %>%
  group_by(eDWID) %>%
  summarise(largest_gap = max(diff(ReviewYear)) - 1)
largest_gap <- data.frame(table(largest_gap$largest_gap))
colnames(largest_gap) <- c("Years between CFRD+", "n")
```

There were a total of `r sum(largest_gap$n)` participants who had a diagnosis of CFRD+ followed by at least one CFRD-, followed again by a CFRD+ diagnosis. For each person, we calculated the largest gap between CFRD+ diagnoses. A total of `r sum(largest_gap$n[as.numeric(largest_gap[,1])>=3])` people had a gap $\geq$ 3 years.

```{r}
kable(largest_gap)
```

>>>>

# Survival analysis - first CFRD diagnosis

@@
